---
title: "index"
subtitle: "Link to Data Source https://usa.ipums.org/usa/"
author: "Benjamin Reese and Sarah Hayes"
format: html
---

```{r}
library(readr)
acs <- read_csv("usa_00003.csv")
library(tidyverse)
```
## Figure One
Figure one examines the income distribution by racial category. As the figure shows, all racial groups have a higher distribution of incomes closer to the poverty line. However, American Indians (Indigenous) have a larger amount of people closer to the poverty line, whereas for Asian Americans, they have the highest distribution of incomes. 
```{r}
acs %>%
  filter(INCTOT < 200000 & INCTOT !=0 ) %>%
  mutate(Race = case_when( ## creating new race variable
    RACE == 1 ~ "White",
    RACE == 2 ~ "Black",
    RACE == 3 ~ "American Indian",
    RACE == 4 | RACE == 5 | RACE == 6 ~ "Asian or P I",
    RACE > 6 ~ "Other"
    )) %>%
  ggplot(aes(Race, INCTOT)) +
  geom_violin(col="blue", fill="light blue") +
  theme_minimal() +
  labs(x=NULL, y="Income",
          title = "The Distribution of Income Is Largely Consistent Between Races",
          subtitle = "Income by Race for Americans",
          caption = "Data Source: IPUMS 2019 ACS")
```

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this 

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
